---
title: "Boostrapping Exercise"
output: github_document
---


### Dependent Variable: support_index
## Survey Questions

Support index is an index that was created from 4 questions about respondent's view on free senior high school.

---

```{r}
# Loading the data
library(haven)
clean_data <- read_dta("~/GitHub/ppol1802/clean_data.dta")


# Convert labeled variables back to factors
clean_data <- as_factor(clean_data)


```{r}


# Load necessary libraries
library(boot)
library(ggplot2)

# Define bootstrap function for mean
boot_mean <- function(data, indices) {
  return(mean(data[indices], na.rm = TRUE)) 
}

# Define bootstrap function for median
boot_median <- function(data, indices) {
  return(median(data[indices], na.rm = TRUE))  
}

# Run bootstrap with 1000 replications
set.seed(123)  # For reproducibility
boot_mean_result <- boot(clean_data$support_index, boot_mean, R = 5)
boot_median_result <- boot(clean_data$support_index, boot_median, R = 5)

# Compute bootstrapped mean and median from resamples
boot_mean_value <- mean(boot_mean_result$t)   
boot_median_value <- mean(boot_median_result$t) 

# Compute observed mean and median
observed_mean <- mean(clean_data$support_index, na.rm = TRUE)
observed_median <- median(clean_data$support_index, na.rm = TRUE)

```



```{r}
# Create a data frame for bootstrapped and observed values
mean_median_df <- data.frame(
  Statistic = c("Mean", "Median"),
  Type = c("Bootstrapped", "Bootstrapped", "Observed", "Observed"),
  Value = c(boot_mean_value, boot_median_value, observed_mean, observed_median)
)

```



```{r}
ggplot(mean_median_df, aes(x = Statistic, y = Value, fill = Type)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.5), color = "black") +
  geom_text(aes(label = round(Value, 3)), position = position_dodge(width = 0.5), vjust = -0.5, size = 5) + 
  labs(title = "Bootstrapped vs. Observed Mean and Median of Support Index",
       x = "Statistic",
       y = "Value") +
  theme_minimal() +
  scale_fill_manual(values = c("Bootstrapped" = "lightblue", "Observed" = "lightgreen"))

```


