---
title: "Codebook"
output: github_document
---

```{r}
# Loading the data
library(haven)
clean_data <- read_dta("~/GitHub/ppol1802/clean_data.dta")


# Convert labeled variables back to factors
clean_data <- as_factor(clean_data)

```

## Data Description
This dataset is a description of Ghanaians view on the Free Senior High School Policy implemented by the New PatrioTic Party in 2017. It also comprises of demographic variables and their partisanship stand.I extracted them from from the Afrobarometer data round 8.


## Variable description

### Dependent Variables: 
## Survey Variables and Questions

### I. Policy Support (Binary)
- **1** = Agree with Statement 1  
- **0** = Agree with Statement 2  

**Question asked in the survey:**  
*"Which of the following statements is closest to your view? Choose Statement 1 or Statement 2."*

- **Statement 1:** *Government should have ensured that all the necessary structures for free SHS are in place before its implementation.*  
- **Statement 2:** *Government took the right decision to start the implementation of the free SHS policy and address the challenges as they arise.*  

---

### II. Opportunity (Categorical)
- **1** = Strongly Disagree  
- **2** = Disagree  
- **3** = Neither Agree nor Disagree  
- **4** = Agree  
- **5** = Strongly Agree  

**Question asked in the survey:**  
*"SHS policy has created opportunity for the poor to obtain education."*

---

### III. Support (Categorical)
- **1** = Strongly Disagree  
- **2** = Disagree  
- **3** = Neither Agree nor Disagree  
- **4** = Agree  
- **5** = Strongly Agree  

**Question asked in the survey:**  
*"It is better to have free SHS even if it leads to an increase in the number of educated citizens who cannot find a job."*

---

### IV. Target (Categorical)
- **1** = Strongly Disagree  
- **2** = Disagree  
- **3** = Neither Agree nor Disagree  
- **4** = Agree  
- **5** = Strongly Agree  

**Question asked in the survey:**  
*"In implementing the free SHS policy, the government should have targeted only the poor who otherwise would not be able to pay for secondary education."*

V)Index 

## Main Independent Variable

### 1. Political Affiliation and Political Party
- **1** = Yes (Affiliated with NPP)  
- **0** = No (Affiliated with NDC)  

**Question asked in the survey:**  
*"Which political party do you feel close to?"*


## Control Variables

Age (Continuous)
Household Size (continuous)
Sex (Binary)
Religion (Binary)
Education (Categorical)
Employment (Categorical)


```{r}
#Summary Statistics
library(dplyr)

clean_data%>% 
  select(opportunity, support, target, policy_support, support_index, age, hh_size, sex, religion, education, employment, urban, present_cond, asset, trust_index, corruption) %>% 
  summary()
```


```{r}
# Convert to data frame explicitly (if needed)
data <- as.data.frame(clean_data)
```


```{r}
library(ggplot2)
library(dplyr)
library(tidyr)


clean_data <- clean_data %>%
  mutate(age = as.numeric(as.character(age)))  # Convert factor to numeric

# Convert data to long format for ggplot
clean_data_long <- clean_data %>%
  select(age, hh_size) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value")

# Plot histograms
ggplot(clean_data_long, aes(x = Value, fill = Variable)) +
  geom_histogram(position = "dodge", bins = 20, alpha = 0.6, color = "black") +
  facet_wrap(~Variable, scales = "free_x") +
  labs(title = "Histograms of Age and Household Size", x = "Value", y = "Frequency") +
  theme_minimal()
```



```{r}
library(ggplot2)
library(dplyr)
library(stringr)

# Remove numbers and drop NAs from REGION names
clean_data <- clean_data %>%
  mutate(REGION = str_remove(REGION, "^[0-9]+\\.\\s*")) %>%  # Removes leading numbers & dots
  filter(!is.na(partisanship))  # Drops NAs from REGION

# Now, plot again with clean labels
ggplot(clean_data, aes(y = REGION, fill = as.factor(partisanship))) +
  geom_bar(position = "fill") +
  scale_x_continuous(labels = scales::percent) +
  labs(title = "Partisanship Distribution by Region",
       y = "Region", x = "Percentage",
       fill = "Partisanship") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))  # Center the title

```


```{r}
library(ggplot2)
library(dplyr)

# Convert 'age' to numeric (if it's stored as character or factor)
data_clean <- clean_data %>%
  mutate(age = as.numeric(as.character(age)))  # Ensures it's numeric

# Bin Age into Groups (5-year intervals)
data_clean <- clean_data %>%
  mutate(age_group = cut(age, 
                         breaks = seq(15, 90, by = 5),  # 5-year bins
                         labels = paste(seq(15, 85, by = 5), "-", seq(19, 89, by = 5)),
                         right = FALSE))

# Ensure Partisanship is a Factor
data_clean$partisanship <- as.factor(data_clean$partisanship)

# Create Density Plot
ggplot(data_clean, aes(x = age, fill = partisanship)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(breaks = seq(15, 90, by = 5)) +  # Reduce axis clutter
  labs(title = "Age Distribution by Partisanship",
       x = "Age Group", y = "Density",
       fill = "Partisanship") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate labels

```


```{r}
library(ggplot2)
library(dplyr)

# Remove NAs from the Political Party variable
data_clean <- data_clean %>%
  filter(!is.na(pol_party))  # Drop missing values

# Create Boxplot Without NAs
ggplot(data_clean, aes(x = as.factor(pol_party), y = trust_index, fill = as.factor(pol_party))) +
  geom_boxplot() +
  labs(title = "Trust in Institutions by Political Party",
       x = "Political Party", y = "Trust Index",
       fill = "Party") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))  # Center the title only


```

```{r}

library(ggplot2)
library(dplyr)

# Remove NAs from Policy Support variable
data_clean <- data_clean %>%
  filter(!is.na(policy_support))  # Drop missing values

# Create Stacked Bar Chart Without NAs
ggplot(data_clean, aes(x = as.factor(policy_support), fill = as.factor(partisanship))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Policy Support by Partisanship",
       x = "Policy Support Level", y = "Percentage",
       fill = "Partisanship") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))  # Center the title only


```

