---
title: "Codebook"
output: github_document
---

```{r}
# Loading the data
library(haven)
clean_data <- read_dta("~/GitHub/ppol1802/clean_data.dta")


# Convert labeled variables back to factors
clean_data <- as_factor(clean_data)

```

## Data Description
This dataset is a description of Ghanaians view on the Free Senior High School Policy implemented by the New PatrioTic Party in 2017. It also comprises of demographic variables and their partisanship stand.These are all pulled from the Afrobarometer data round 8.


## Variable description
### Main Dependent Variable: policy_support (Binary) Agree with statement 1 is 1 and Agree with Statement 2 is 0.
Question asked in the survey: 

Statement 1:Government should have ensured that all the necessary structures for free SHS are in place before its implementation. 

Statement 2: Statement 2: Government took the right decision to start the implementation of the free SHS policy and address the challenges as they arise.   


## Main independent variables: 
Independent Variable 1 (Political Affiliation - Binary). Question:  Do you feel close to any particular political party?

Yes is 1 and No otherwise.


Independent Variable 2 (Politically Affiliation and Political Party)

1 is Yes (NPP) and 0 is No (NDC)

## Control Variables

Age (Continuous)
Household Size (continuous)
Sex (Binary)
Religion (Binary)
Education (Categorical)
Employment (Categorical)


```{r}
#Summary Statistics
library(dplyr)

clean_data%>% 
  select(policy_support, age, hh_size, sex, religion, education, employment) %>% 
  summary()
```



```{r}
library(ggplot2)
library(dplyr)
library(tidyr)


clean_data <- clean_data %>%
  mutate(age = as.numeric(as.character(age)))  # Convert factor to numeric

# Convert data to long format for ggplot
clean_data_long <- clean_data %>%
  select(age, hh_size) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value")

# Plot histograms
ggplot(clean_data_long, aes(x = Value, fill = Variable)) +
  geom_histogram(position = "dodge", bins = 20, alpha = 0.6, color = "black") +
  facet_wrap(~Variable, scales = "free_x") +
  labs(title = "Histograms of Age and Household Size", x = "Value", y = "Frequency") +
  theme_minimal()
```



